### 2024-09-05 texts/sup_materials.tex
update sup materials, also include tex version

% Created 2024-09-05 do 11:15
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

\usepackage[hyphens]{url}                
\usepackage{hyperref}
\usepackage[hyphenbreaks]{breakurl}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[section]{placeins}
\usepackage{dirtree}
\usepackage{siunitx}
\usepackage{afterpage}
\usepackage{pdflscape}
\usepackage{svg}

\usepackage{booktabs}
\usepackage{dcolumn}

\usepackage{bibentry}

\sisetup{detect-all}

\sloppy
\usepackage{scalerel,stackengine}

\stackMath
\newcommand\reallywidehat[1]{%
\savestack{\tmpbox}{\stretchto{%
  \scaleto{%
    \scalerel*[\widthof{\ensuremath{#1}}]{\kern-.6pt\bigwedge\kern-.6pt}%
    {\rule[-\textheight/2]{1ex}{\textheight}}%WIDTH-LIMITED BIG WEDGE
  }{\textheight}% 
}{0.5ex}}%
\stackon[1pt]{#1}{\tmpbox}%
}

\usepackage{caption}
\usepackage[draft]{todonotes}

\captionsetup{skip=0pt}
\usepackage[utf8]{inputenc}
\usepackage[style=apa, backend=biber]{biblatex} 
\usepackage[english, american]{babel}
\DeclareLanguageMapping{american}{american-apa}
\DeclareFieldFormat{apacase}{#1}

\usepackage[T1]{fontenc}
\usepackage{csquotes}

\addbibresource{/home/johannes/Dropbox/references.bib}
\addbibresource{/home/johannes/Dropbox/references2.bib}

\usepackage{floatrow}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{colortbl}

\lstset{
  language=R,                    
  basicstyle=\footnotesize,      
  numbers=left,                  
  numberstyle=\tiny\color{gray}, 
  stepnumber=1,                  
  numbersep=5pt,                 
  backgroundcolor=\color{white}, 
  showspaces=false,              
  showstringspaces=false,        
  showtabs=false,                
  frame=single,                  
  rulecolor=\color{black},       
  tabsize=2,                     
  captionpos=b,                  
  breaklines=true,               
  breakatwhitespace=false,       
  title=\lstname,                
  keywordstyle=\color{red},     
  commentstyle=\color{blue},  
  stringstyle=\color{violet},     
  escapeinside={\%*}{*)},        
  morekeywords={*,...}           
}

% \usepackage{crimson}
% \usepackage{microtype}

% \usepackage{helvet}
% \renewcommand{\familydefault}{\sfdefault}

\usepackage{tgtermes} % times font

\usepackage{fancyhdr}
\usepackage{setspace}
\onehalfspacing
\usepackage{longtable}
\usepackage{subfig}
% \usepackage[a4paper, total={18cm, 24cm}]{geometry}
\usepackage[a4paper, margin=2.5cm]{geometry}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\maketitle}{}

\usepackage{enumitem}
\setlist[itemize]{topsep=0pt,itemsep=0pt,parsep=0pt,partopsep=0pt}

\usepackage{multicol}
\setlength\multicolsep{0pt}

\newlist{propertyList}{itemize}{1}
\setlist[propertyList]{
  label=\textbullet,
  noitemsep,
  leftmargin=10pt,
  before=\begin{multicols}{3},
  after=\end{multicols}
  }

% \cfoot {Johannes Aengenheyster}
\rfoot {\thepage}

\listfiles

\setlength{\parindent}{1.2cm}
\author{Johannes }
\date{\today}
\title{}
\hypersetup{
 pdfauthor={Johannes },
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.4 (Org mode 9.7.9)}, 
 pdflang={English}}
\begin{document}

\section*{Supplementary online materials}

\subsection*{Alternative competition specifications}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Thu Sep  5 09:54:41 2024
\begin{table}[ht]
\centering
\begin{tabular}{p{0mm}lD{)}{)}{8)3}D{)}{)}{8)3}D{)}{)}{8)3}D{)}{)}{8)3}D{)}{)}{8)3}}
  \hline 
 \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{r\_pop4} & \multicolumn{1}{l}{r\_pmdens1} & \multicolumn{1}{l}{r\_pmdens2} & \multicolumn{1}{l}{r\_audience1} & \multicolumn{1}{l}{r\_audience\_log1}\\ 
 \hline
  \multicolumn{7}{l}{\textbf{Founder}} \\ 
 & Gender - Female & -.04 \; (0.34) & .00 \; (0.33) & -.05 \; (0.33) & -.00 \; (0.33) & -.11 \; (0.33) \\ 
   & Gender - Couple & -.14 \; (0.32) & -.22 \; (0.32) & -.18 \; (0.32) & -.21 \; (0.32) & -.23 \; (0.32) \\ 
   & Founder died & .28 \; (0.43) & .19 \; (0.43) & .25 \; (0.43) & .18 \; (0.43) & .08 \; (0.43) \\ 
   \multicolumn{7}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Foundation & .20 \; (0.41) & .36 \; (0.40) & .33 \; (0.40) & .37 \; (0.40) & .44 \; (0.40) \\ 
   & Self-Identification - Collection & -.45 \; (0.54) & -.09 \; (0.50) & -.20 \; (0.51) & -.05 \; (0.51) & .05 \; (0.51) \\ 
   & Self-Identification - Other & 1.09 \; (0.30)^{***} & 1.03 \; (0.30)^{***} & 1.12 \; (0.31)^{***} & 1.05 \; (0.30)^{***} & 1.13 \; (0.30)^{***} \\ 
   & AN Ranking - Included & .19 \; (0.34) & .24 \; (0.34) & .17 \; (0.35) & .23 \; (0.34) & .18 \; (0.34) \\ 
   & AN Ranking - Dropped & .46 \; (0.50) & .45 \; (0.50) & .31 \; (0.50) & .45 \; (0.49) & .47 \; (0.49) \\ 
   & Founder name in Museum name & .05 \; (0.28) & .10 \; (0.28) & .14 \; (0.28) & .10 \; (0.28) & .07 \; (0.28) \\ 
   \multicolumn{7}{l}{\textbf{Environment}} \\ 
 & PM density (country) & .65 \; (0.34) & .50 \; (0.32) & .41 \; (0.34) & .55 \; (0.33) & .73 \; (0.31)^{*} \\ 
   & PM density$^{2}$ (country) & -.03 \; (0.02) & -.02 \; (0.02) & -.02 \; (0.02) & -.02 \; (0.02) & -.03 \; (0.03) \\ 
   & Pop. (millions) within 10km & .27 \; (0.09)^{**} &  &  &  &  \\ 
   & Nbr PM within 10km & .26 \; (0.08)^{***} &  &  &  &  \\ 
   & PM density (10km) &  & -.00 \; (0.03) &  &  &  \\ 
   & PM density (10km, log) &  &  & .29 \; (0.15) &  &  \\ 
   & Nbr PM (10km) * Pop (10km) & -.09 \; (0.03)^{**} &  &  &  &  \\ 
   & Local Audience per PM &  &  &  & .10 \; (0.14) &  \\ 
   & Local Audience per PM (log) &  &  &  &  & .28 \; (0.09)^{**} \\ 
   & Great Recession (2008/09) & -1.59 \; (1.01) & -1.65 \; (1.01) & -1.64 \; (1.01) & -1.65 \; (1.01) & -1.63 \; (1.01) \\ 
   & Covid Pandemic (2020/21) & .42 \; (0.29) & .44 \; (0.29) & .46 \; (0.29) & .43 \; (0.29) & .42 \; (0.29) \\ 
   \hline
 & Museums & 503 & 503 & 503 & 503 & 503 \\ 
   & Museum-years & 6096 & 6096 & 6096 & 6096 & 6096 \\ 
   & Closures & 68 & 68 & 68 & 68 & 68 \\ 
   & log. Likelihood & -339.68 & -347.93 & -346.18 & -347.72 & -342.93 \\ 
   & AIC & 711.35 & 723.85 & 720.35 & 723.44 & 713.86 \\ 
   & BIC & 746.87 & 754.93 & 751.42 & 754.51 & 744.93 \\ 
   \hline 
 \multicolumn{7}{l}{\footnotesize{standard errors in parantheses.\textsuperscript{***}p $<$ 0.001;\textsuperscript{**}p $<$ 0.01;\textsuperscript{*}p $<$ 0.05.}}
\end{tabular}
\caption{Alternative specification of environment} 
\label{tbl:t_reg_coxph_env}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics[width=18cm]{../figures/p_pred_heatmap_alt.pdf}
\caption{\label{fig:p_pred_heatmap_alt}Prediction Heatmap of Alternative model specifications}
\end{figure}

Alternative Specifications of competition exist.
First, it is possible to argue that the primary mechanism through which private museums compete with each other is by limiting each other's audience.
This specification leads to the construction of the Audience variable, which describes the number of people in the 10 km surrounding by the number of private museums present.
Under this assumption, the situation of a private museum which is alone in an environment of 1 million people is equivalent to one in an environment of 3 million people with 2 other private museums, as each museums is "assigned" a pool of 1 million people.
Secondly, one might argue that the competition between private museums is scaled by the amount of people, which corresponds to specifying competition as the number of other private museums per (million) population.
In this case, the absence of other private museums assumes absence of competition, regardless of the audience size.

These while it would be possible for these different specifications to lead to similar interpretations, in practice they diverge substantially (table \ref{tbl:t_reg_coxph_env}).
Using Private museum density (model 2 and 3 for linear and log-transformed audience, figure \ref{fig:p_pred_heatmap_alt} bottom left), the general conclusion would be that (logged) competition increases mortality as predicted closing chance is highest in areas of low population and large private museum populations.
However, using audience (models 4 and 5, figure \ref{fig:p_pred_heatmap_alt} bottom right) leads to the opposite conclusion, as higher audience is associated with higher mortality; predicted closing chance is highest for areas of low numbers of private museums and large populations.
While it is beyond the scope of the paper to explore the origin of these diverging perspectives in detail, it is worth pointing out that the main specification of private interacted private museum and population counts captures both of these relationships, which arguably makes it preferable for its encompassingness.
\subsection*{Alternative sample specifications}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Thu Sep  5 11:13:39 2024
\begin{table}[ht]
\centering
\begin{tabular}{p{0mm}lD{)}{)}{8)3}D{)}{)}{8)3}D{)}{)}{8)3}D{)}{)}{8)3}}
  \hline 
 \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{r\_pop4} & \multicolumn{1}{l}{r\_wocryside} & \multicolumn{1}{l}{r\_onlycryside} & \multicolumn{1}{l}{r\_pop42}\\ 
 \hline
  \multicolumn{6}{l}{\textbf{Founder}} \\ 
 & Gender - Female & -.04 \; (0.34) & .21 \; (0.44) & -1.07 \; (0.78) & -.01 \; (0.34) \\ 
   & Gender - Couple & -.14 \; (0.32) & .07 \; (0.52) & -.27 \; (0.46) & -.12 \; (0.32) \\ 
   & Founder died & .28 \; (0.43) & -.19 \; (0.77) & .58 \; (0.59) & .29 \; (0.43) \\ 
   \multicolumn{6}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Foundation & .20 \; (0.41) & 1.02 \; (0.65) & .31 \; (0.58) & .18 \; (0.41) \\ 
   & Self-Identification - Collection & -.45 \; (0.54) & .95 \; (0.71) & -18.23 \; (4781.88) & -.50 \; (0.55) \\ 
   & Self-Identification - Other & 1.09 \; (0.30)^{***} & 1.95 \; (0.51)^{***} & .67 \; (0.46) & 1.09 \; (0.31)^{***} \\ 
   & AN Ranking - Included & .19 \; (0.34) & -.32 \; (0.48) & .55 \; (0.58) & .18 \; (0.34) \\ 
   & AN Ranking - Dropped & .46 \; (0.50) & .18 \; (0.73) & -.06 \; (0.85) & .47 \; (0.50) \\ 
   & Founder name in Museum name & .05 \; (0.28) & -.30 \; (0.42) & .24 \; (0.44) & .04 \; (0.28) \\ 
   \multicolumn{6}{l}{\textbf{Environment}} \\ 
 & PM density (country) & .65 \; (0.34) & -1.16 \; (2.52) & .88 \; (0.44)^{*} & .67 \; (0.34)^{*} \\ 
   & PM density$^{2}$ (country) & -.03 \; (0.02) & 1.33 \; (2.38) & -.04 \; (0.02) & -.03 \; (0.02) \\ 
   & Pop. (millions) within 10km & .27 \; (0.09)^{**} & .17 \; (0.17) & 1.69 \; (0.81)^{*} & .20 \; (0.12) \\ 
   & Nbr PM within 10km & .26 \; (0.08)^{***} & .19 \; (0.10) & .64 \; (1.06) & .22 \; (0.20) \\ 
   & Nbr PM$^{2}$ within 10km &  &  &  & .01 \; (0.02) \\ 
   & Nbr PM (10km) * Pop (10km) & -.09 \; (0.03)^{**} & -.07 \; (0.04)^{*} & -.54 \; (1.83) & -.02 \; (0.09) \\ 
   & Nbr PM$^{2}$ (10km) * Pop (10km) &  &  &  & -.01 \; (0.01) \\ 
   & Great Recession (2008/09) & -1.59 \; (1.01) & -18.42 \; (5245.29) & -.75 \; (1.04) & -1.59 \; (1.01) \\ 
   & Covid Pandemic (2020/21) & .42 \; (0.29) & .30 \; (0.41) & .46 \; (0.47) & .45 \; (0.29) \\ 
   \hline
 & Museums & 503 & 219 & 283 & 503 \\ 
   & Museum-years & 6096 & 2265 & 3311 & 6096 \\ 
   & Closures & 68 & 35 & 28 & 68 \\ 
   & log. Likelihood & -339.68 & -139.36 & -116.47 & -338.79 \\ 
   & AIC & 711.35 & 310.72 & 264.94 & 713.59 \\ 
   & BIC & 746.87 & 335.60 & 286.25 & 753.54 \\ 
   \hline 
 \multicolumn{6}{l}{\footnotesize{standard errors in parantheses.\textsuperscript{***}p $<$ 0.001;\textsuperscript{**}p $<$ 0.01;\textsuperscript{*}p $<$ 0.05.}}
\end{tabular}
\caption{Cox PH models of different local density specifications} 
\label{tbl:t_reg_coxph_density}
\end{table}

To investigate the robustness of the coefficients across different values, table \ref{tbl:t_reg_coxph_density} shows how coefficients according to sample restrictions. 
In model 1, only museums are included which are located in areas with the highest value combination of number and private museums and population, i.e. museums located in an environment of less than 2 million people and 0 or 1 other private museums (the "countryside").
In model 2, the "countryside" is excluded.
However, both these models are not reliable, as they do not properly converge (indicated by extremely large coefficients in both models).
\end{document}